PAMELA: an annotation-based Java Modelling Framework

### 2020-07-03 4-ApproachOverview.tex
Ajout paragraphe PAMELA use process

Coupling model and code into the same artefact open new ways of programming. The classical way relies on \emph{programmers} that produce code reusing pre-existing modeling concepts. These concepts are implemented by \emph{modelers} that provides the right annotations the programmers use. This is, for instance, the process followed by JEE developers reusing JEE specific annotations. The evolution rhythm between models and code is low. This programming way is still possible with PAMELA, but we allow the ability to reach a high evolution rhythm when the programmer becomes also the modeler. In fact, when a pattern, an abstraction, a generalization is identified by the programmer, s/he can use PAMELA to develop and capitalize on this abstraction by increasing PAMELA metamodel. Our experience shows that, in that case, introducing and reusing new concepts (1) reduce the size of the code, (2) reduce the risk of errors and (3) improve the code structure. The cycle of development between the model and the code can then be drastically reduced, leading to what we call \emph{continuous modeling}.

The code size is reduced because abstractions factorize the recognize pattern and the previous code is replaced by the use of the abstraction at the right place. This also reduce the risk since the previous code is now generated by the PAMELA framework with all the required verification. And finally, the code structure is improved since it matches the way the programmer conceptualizes (models) her/his code.

The developed metamodels are implemented by annotations that relies on Java/JVM entities and mechanisms. They include consistency checking that constrains their use and help the programmer. We have experimented their use with setter/getter to define POJO entities, with traits to implement multiple inheritance or roles and rules to set security rules on classes.

>>>>

Coupling model and code into the same artefact open new ways of programming. The classical way relies on \emph{programmers} that produce code reusing pre-existing modeling concepts. These concepts are implemented by \emph{modelers} that provides the right annotations the programmers use. This is, for instance, the process followed by JEE developers reusing JEE specific annotations. The evolution rhythm between models and code is low. This programming way is still possible with PAMELA, but we allow the ability to reach a high evolution rhythm when the programmer becomes also the modeler. In fact, when a pattern, an abstraction, a generalization is identified by the programmer, s/he can use PAMELA to develop and capitalize on this abstraction by increasing PAMELA metamodel. 
The developed metamodels are implemented by annotations that relies on Java/JVM entities and mechanisms. They include consistency checking that constrains their use and help the programmer. We have experimented their use with setter/getter to define POJO entities, with traits to implement multiple inheritance or roles and rules to set security rules on classes.

Our experience shows that introducing and reusing new concepts (1) reduce the size of the code, (2) reduce the risk of errors and (3) improve the code structure. The cycle of development between the model and the code can then be drastically reduced, leading to what we call \emph{continuous modeling}.

The code size is reduced because abstractions factorized the recognize pattern and the previous code is replaced by the use of the abstraction at the right place. This also reduces the risk since the previous code is now generated by the PAMELA framework with all the required checks. And finally, the code structure is improved since it matches the way the programmer conceptualizes (models) her/his code.
